Ecosystem size and complexity dictate riverine biodiversity

### 2021-01-05 manuscript.Rmd
bulk edits

Despite the substantial difference in fish fauna between the study regions, the estimated Bayes factor in favor of the global model explaining $\gamma$ diversity was `r round(fit$gamma$bayes_factor, 1)`; this result strongly supports the consistent effects of ecosystem size and complexity on $\gamma$ diversity across the two biogeographic regions. Further, we found patterns consistent with our theoretical predictions. The estimated $\gamma$ diversity increased with increasing watershed area (ecosystem size) and branching probability (ecosystem complexity) across watersheds. These effects remained significant even after controlling for the potential influences of other environmental factors (precipitation, temperature, and land use). Overall, our statistical analysis provides strong empirical evidence that ecosystem size and complexity jointly, but independently, scale riverine biodiversity at the metacommunity level.

@@

While the prevailing evidence supports the importance of ecosystem size in scaling species diversity, ecosystem complexity - especially branching structure - has not received the attention it deserves. Ecosystems are inherently complex in its spatial structure, providing a physical template that creates a wide spectrum of niche opportunities for living organisms. Hence, our findings should be broadly applicable to many taxa and ecosystems. Generalization of the complexity perspective is a frontier for future research.

@@

## Theoretical branching networks

@@

## Metacommunity model
We simulated metacommunity dynamics in branching river networks following the metacommunity framework proposed by Thompson et al.[@thompsonProcessbasedMetacommunityFramework2020]. In this framework, abiotic environmental conditions (density independent), intra- and interspecific competition (density dependent), and dispersal regulate the realized population growth of a species. Below, we describe details of our metacommunity model.

@@

Smaller values of $\nu$ imply greater costs of having wider niche (i.e., decreased maximum reproductive rate) and reduce the realized reproductive rate $r_{ix}(t)$.

@@

## Simulation
We used `r n_scenario` sets of parameter combinations to produce different landscape (four patterns of habitat heterogeneity) and ecological scenarios (eight patterns of dispersal and competition). To create landscape scenarios with different patterns of habitat heterogeneity, we varied environmental variation at headwaters ($\sigma_h$) and the degree of local environmental noise ($\sigma_l$) separately. Larger values of $\sigma_h$ and/or $\sigma_l$ produce greater variation in environmental conditions among habitat patches. We used two values for each of the SD parameters ($\sigma_h =$ `r unique(dat_sim$sd_env_source)` and $\sigma_l =$ `r unique(dat_sim$sd_env_lon)`), resulting in four sets of parameter combinations. When $\sigma_h > \sigma_l$, having more tributaries in a network exhibits important effects on overall habitat diversity as branching is the primary source of environmental variation. When $\sigma_h \le \sigma_l$, local environmental noise masks environmental variation among tributaries, leading to minimal influences of branching on overall habitat diversity in a network. For ecological scenarios, we varied three ecological parameters that are relevant for dispersal and interspecific competition: dispersal distance ($\theta =$ `r unique(dat_sim$theta)`), dispersal probability ($p_d =$ `r unique(dat_sim$p_dispersal)`), and the maximum value of competition coefficients ($\alpha_{max} =$ `r unique(dat_sim$max_alpha)`). This setup results in eight ecological scenarios with different levels of dispersal and competition.

@@

## Empirical data
We compiled fish data at two biogeographic regions: Hokkaido island, Japan and Midwest, US. For Hokkaido, we used data from the Hokkaido Freshwater Fish Database HFish, monitoring data at protected watersheds, and primary data collected by the authors. Collectively, these fish data encompass more than 7000 sites and covers the entire Hokkaido island. For the Midwest, we assembled fish community data collected by XXX.

We selected data based on the following criterion. First, we reviewed

>>>>

Despite the substantial difference in fish fauna between the study regions, the estimated Bayes factor in favor of the global model was `r round(fit$gamma$bayes_factor, 1)` for $\gamma$ diversity; this result strongly supports the consistent effects of ecosystem size and complexity on $\gamma$ diversity across the two biogeographic regions. Further, we found patterns consistent with our theoretical predictions. The estimated $\gamma$ diversity increased with increasing watershed area (ecosystem size) and branching probability (ecosystem complexity) across watersheds. These effects remained significant even after controlling for the potential influences of other environmental factors (precipitation, temperature, and land use). Overall, our statistical analysis provides strong empirical evidence that ecosystem size and complexity jointly, but independently, scale riverine biodiversity at the metacommunity level.

@@

While the prevailing evidence supports the importance of ecosystem size in scaling species diversity, ecosystem complexity - especially branching structure - has not received the attention it deserves. Ecosystems are inherently complex in its spatial structure, providing a physical template that creates a wide spectrum of niche opportunities for living organisms. Hence, our findings should be broadly applicable to many taxa and ecosystems. Generalization of our findings to other spatially complex ecosystems represents a frontier for future research.

@@

## Theory

### Theoretical branching networks

@@

### Metacommunity model
We simulated metacommunity dynamics in branching river networks following the metacommunity framework proposed by Thompson et al.[@thompsonProcessbasedMetacommunityFramework2020]. In this framework, abiotic environmental conditions (density independent), intra- and interspecific competition (density dependent), and dispersal regulate the realized population growth of a species. Below, we describe details of our metacommunity model.

@@

Smaller values of $\nu$ imply greater costs of having wider niche (i.e., decreased maximum reproductive rate) and reduce the realized reproductive rate $r_{ix}(t)$ near the niche optimum.

@@

### Simulation
We used `r n_scenario` sets of parameter combinations to produce different landscape (four patterns of habitat heterogeneity) and ecological scenarios (eight patterns of dispersal and competition). To create landscape scenarios with different patterns of habitat heterogeneity, we varied environmental variation at headwaters ($\sigma_h$) and the degree of local environmental noise ($\sigma_l$) separately. Larger values of $\sigma_h$ and/or $\sigma_l$ produce greater variation in environmental conditions among habitat patches. We used two values for each of the SD parameters ($\sigma_h =$ `r unique(dat_sim$sd_env_source)` and $\sigma_l =$ `r unique(dat_sim$sd_env_lon)`), resulting in four sets of parameter combinations. When $\sigma_h > \sigma_l$, having more tributaries in a network exhibits important effects on overall habitat diversity as branching is the primary source of environmental variation. When $\sigma_h \le \sigma_l$, local environmental noise masks environmental variation among tributaries, leading to minimal influences of branching on overall habitat diversity in a network. For ecological scenarios, we varied three ecological parameters that are relevant for dispersal and interspecific competition: dispersal distance ($\theta =$ `r unique(dat_sim$theta)`), dispersal probability ($p_d =$ `r unique(dat_sim$p_dispersal)`), and the maximum value of competition coefficients ($\alpha_{max} =$ `r unique(dat_sim$max_alpha)`). This setup results in eight ecological scenarios with different levels of dispersal and competition.

@@

## Empirical data analysis

### Fish community data

We compiled fish data at two biogeographic regions: Hokkaido island, Japan and Midwest, US. For Hokkaido, we used data from the Hokkaido Freshwater Fish Database HFish, monitoring data at protected watersheds, and primary data collected from literature. Collectively, these fish data encompass more than 7000 sites and covers the entire Hokkaido island. For the Midwest, we assembled fish community data collected by XXX. We detailed procedures for initial data screening in SI.

We further selected data based on the following criteria. First, we selected sampling sites with reliable coordinates. Sampling sites were treated as the same site if coordinates are identical at 3 decimal degrees. Second, for sites with multiple site visits (i.e., temporal replicates), we used the latest data at each sampling site to minimize variation in sampling efforts among sites. Lastly, we selected metacommunities (i.e., watersheds) that contain more than 10 sampling sites to reduce statistical uncertainties in estimating $\gamma$ diversity. We considered watersheds as separated from others if the river network flows into one of the following: the ocean, a large lake ($\ge$ 10 km² in the areal area) or a large river that may represent lentic habitats ($\ge$ 5000 km² in watershed area).

We estimated $\alpha$, $\beta$ and $\gamma$ diversity at each watershed. $\gamma$ diversity is the total species richness at each watershed. Since it is impossible to sample all species present in each watershed, we used asymptotic species richness as a proxy for $\gamma$ diversity. Asymptotic species richness is the estimated total species richness in a given area and can be estimated from spatial replicates of local community data within a metacommunity. We estimated asymptotic species richness using sampling sites located in the same watershed (i.e., spatial replicates within a metacommunity). We assessed the completeness of the observed species richness based on sample coverage, which is the proportion of observed species richness to asymptotic species richness. We excluded watersheds whose sample coverage was below 0.9 because the estimates of asymptotic species richness may contain non-negligible statistical uncertainty. $\alpha$ diversity was estimated by taking an average of local species richness at each site. $\beta$ diversity was estimated by dividing the estimated $\gamma$ diversity by $\alpha$ diversity.

After the data selection procedure, we obtained diversity estimates from `r nrow(wsd_subset[[1]]) + nrow(wsd_subset[[2]])` watersheds (Hokkaido: `r nrow(wsd_subset[[1]])`, Midwest: `r nrow(wsd_subset[[2]])`), which encompassed a total of `r nrow(point_subset[[1]]) + nrow(point_subset[[2]])` sites (Hokkaido: `r nrow(point_subset[[1]])`, Midwest: `r nrow(point_subset[[2]])`).

### Environmental data

We estimated watershed characteristics (watershed area and branching probability), climatic variables (annual mean air temperature and cumulative precipitation), and land use (% forest). We delineated watershed polygons and riverlines using Hydrologically Adjusted Elevations of MERIT Hydro[@yamazakiMERITHydroHighResolution2019] with ArcMap 10.7 and SAGA 6.3.0 combined with an R package 'RSAGA'[@brenningRSAGASAGAGeoprocessing2018]. Watershed area was estimated based on the delineated watershed polygons using functions in an R package 'sf'[@pebesmaSimpleFeaturesStandardized2018]. Branching probability was estimated by fitting exponential distributions to histograms of branch length (unit: kilometer, river segment between successive confluences or a confluence and the outlet/upstream terminal) as $branch~length \sim Exp(\lambda)$. A derived statistical quantity $1 − e^{−\lambda}$ corresponds to the theoretical branching probability $P_b$. For each watershed, we estimated the spatial averages of air temperature and cumulative precipitation at a 1-km resolution, which were extracted from WorldClim 1.4[@hijmansVeryHighResolution2005] via R package 'raster'. We calculated % forest for each watershed using land use data in 2015 from Copernicus Global Land Service (100-m resolution)[@marcelbuchhornCopernicusGlobalLand2020].

### Statistical analysis
